Rational Monte Carlo Sampling in Trees

### 2012-01-11 paper/rmcts/paper.tex
paper in progress

The {\bf two-stage} Monte Carlo tree search sampling
scheme selects an action at the current root node according to an
algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
UQB, and then selects actions according to UCB. Such two-stage
outperforms the UCT sampling scheme that selects actions based on
UCB only.

Such sampling scheme is significantly less sensitive to the tuning of
the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since the
conflict \cite{Bubeck.pure} between the need for a larger value of
$C_p$ on the first step (simple regret) and a smaller value for the
rest of the rollout (cumulative regret) is resolved. In fact, a
sampling scheme that uses UCB at all steps but a larger value of $C_p$
for the first step than for the rest of the steps, outperforms
UCT. The pseudocode of the two-stage rollout is in
Algorithm~\ref{alg:two-stage-mcts}.

>>>>

The {\bf two-stage} Monte Carlo tree search sampling
scheme selects an action at the current root node according to an
algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
UQB, and then selects actions according to UCB. This scheme
outperforms the UCT sampling scheme that selects actions based on
UCB only.

Additionaly, such two-stage sampling scheme is significantly less
sensitive to the tuning of the exploration factor ($C_p$ in
\cite{Kocsis.uct}) of UCT, since the conflict \cite{Bubeck.pure}
between the need for a larger value of $C_p$ on the first step (simple
regret) and a smaller value for the rest of the rollout (cumulative
regret) is resolved. In fact, a sampling scheme that uses UCB at all
steps but a larger value of $C_p$ for the first step than for the rest
of the steps, outperforms UCT. The pseudocode of the two-stage rollout
is in Algorithm~\ref{alg:two-stage-mcts}.
